### 2018-07-25 express2018/contraction.tex
Fixed p.5 l.22: instead of \subseteq write \subset or \subsetneq. #27
Minor typesetting optimizations
Make the definition of "contracts" consistent with bisimilarity definition.

\begin{definition}[bisimulation  contraction]
\label{d:BisCon}
A process relation ${\R}$ 
 is a {\em  bisimulation  contraction}  if, whenever
 $P\RR Q$, %for all $\mu$

\begin{enumerate}

\item   $P \arr\mu P'$ implies there is $Q'$ such that $Q \arcap \mu
  Q'$
 and $P' \RR Q'$;

\item 
    $Q \arr\mu Q'$   implies there is $P'$ such that $P \Arcap \mu
 P'$ and $P' 
\wb Q'$.
\end{enumerate}
\emph{Bisimilarity  contraction}, written $\mcontrBIS$, is the union
of all bisimulation contractions. 
\end{definition}
% We say that $P$  {\emph{bisimilarly  contracts} $Q$, written
% $Q \mcontrBIS Q$.
% We write $\mexpaBIS$ for the inverse of $\mcontrBIS$.

@@

Bisimilarity  contraction is coarser than 
 the \emph{expansion  relation} 
$\expa$ \cite{sangiorgi2015equations} %, arun1992efficiency}.
This is a
preorder widely used in proof techniques for bisimilarity and that 
intuitively refines bisimilarity by 
 formalising the idea of `efficiency' between processes.
Clause (1) is the same in the two
preorders. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \expa Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the  possibility of
being at least as efficient as $P$.

@@

\begin{example}
\label{exa:contr}
We have %\mcontrBIS a + \tau^n . a $
 $ a \not  \mcontrBIS \tau. a$. However,
$a+ \tau . a \mcontrBIS a$, as well as its converse, 
$  a \mcontrBIS a +
\tau . a $. Indeed, if $P \wb Q$ then 
$  P  \mcontrBIS P +Q$. The last two relations do not hold with 
$\expa$, which explains the strictness of the inclusion
 ${\expa} \subseteq {\mcontrBIS}$. 
% The inclusion is strict: for instance
% $a+ \tau . a \mcontrBIS a$, where $\mcontrBIS$ cannot be replaced by
%  $\contr$. Also the converse of  $a+ \tau . a \mcontrBIS a$ holds, namely
% $  a \mcontrBIS a +
% \tau . a $. However, we have %\mcontrBIS a + \tau^n . a $
%  $ a \not  \mcontrBIS \tau. a$
\end{example}

@@

As  bisimilarity  and expansion, contraction is preserved by all operators but sum.

>>>>

\begin{definition}%[bisimulation contraction]
\label{d:BisCon}
A process relation ${\R}$ 
 is a {\em  bisimulation  contraction}  if, whenever
 $P\RR Q$, %for all $\mu$

\begin{enumerate}
\item   $P \arr\mu P'$ implies there is $Q'$ such that $Q \arcap \mu
  Q'$
 and $P' \RR Q'$;

\item 
    $Q \arr\mu Q'$   implies there is $P'$ such that $P \Arcap \mu
 P'$ and $P' 
\wb Q'$.
\end{enumerate}
\emph{Bisimilarity  contraction}, written as $P \mcontrBIS Q$ ($P$
\emph{contracts to} $Q$), if $P\ \R\ Q$ for some contraction $\R$.
\end{definition}
% We say that $P$  {\emph{bisimilarly  contracts} $Q$, written
% $Q \mcontrBIS Q$.
% We write $\mexpaBIS$ for the inverse of $\mcontrBIS$.

@@

Bisimilarity  contraction is coarser than 
 the \emph{expansion relation} 
$\expa$ \cite{arun1992efficiency,sangiorgi2015equations}.
This is a
preorder widely used in proof techniques for bisimilarity and that 
intuitively refines bisimilarity by 
 formalising the idea of `efficiency' between processes.
Clause (1) is the same in the two
preorders. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \expa Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the  possibility of
being at least as efficient as $P$.

@@

\begin{example}
\label{exa:contr}
We have %\mcontrBIS a + \tau^n . a $
 $ a \not  \mcontrBIS \tau. a$. However,
$a+ \tau . a \mcontrBIS a$, as well as its converse, 
$  a \mcontrBIS a +
\tau . a $. Indeed, if $P \wb Q$ then 
$  P  \mcontrBIS P +Q$. The last two relations do not hold with 
$\expa$, which explains the strictness of the inclusion
 ${\expa} \subset {\mcontrBIS}$. 
% The inclusion is strict: for instance
% $a+ \tau . a \mcontrBIS a$, where $\mcontrBIS$ cannot be replaced by
%  $\contr$. Also the converse of  $a+ \tau . a \mcontrBIS a$ holds, namely
% $  a \mcontrBIS a +
% \tau . a $. However, we have %\mcontrBIS a + \tau^n . a $
%  $ a \not  \mcontrBIS \tau. a$
\end{example}

@@

\hl{Like} (weak) bisimilarity and expansion, contraction is
preserved by all operators but (direct) sum.
